Relational Parametricity for Algebraically Indexed Types

### 2012-06-05 paper/section-general-framework.tex
More updates to the general framework section

  \sigma^*(\forall i\mathord:s.A) \isDefinedAs \forall i\mathord:s.\sigma_{i\mathord:s}^*A
\end{mathpar}
\begin{lemma}
  Let $\sigma : \Delta \Rightarrow \Delta'$ be a simultaneous
  substitution.
  \begin{enumerate}
  \item If $\Delta' \vdash e : s$, then $\Delta \vdash \sigma^*e : s$; and
  \item If $\Delta' \vdash A \isType$, then $\Delta \vdash \sigma^*A
    \isType$.
  \end{enumerate}
\end{lemma}

@@

Much of the power of indexing types by the terms of an algebraic
theory comes from the equations of the theory, and the fact that these
equations are inherited by the types. In the units of measure example,
the well-indexed types $\tyPrim{X}{u_1\cdot u_2}$ and $\tyPrim{X}{u_2
  \cdot u_1}$ are considered equal by the type system, due to the
commutativity of the $-\cdot-$ operation in the theory of abelian
groups.

In our general framework, the equations between types are derived from
a set $\IndexAxiomSet$ of axioms $\Delta \vdash e \stackrel{ax}\equiv
e' : s$ between well-sorted index terms. The constraint that axioms
are well-sorted means that both $\Delta \vdash e : s$ and $\Delta
\vdash e' : s$ for all axioms $(\Delta \vdash e \stackrel{ax}\equiv e'
: s) \in \IndexAxiomSet$.

Given a set $\IndexAxiomSet$ of axioms, we generate the judgmental
equality between index terms $\Delta \vdash e \equiv e' : s$ by the
following rules, which allow us to use substitution instances of
axioms, and ensure that judgmental equality is a congruence relation
with respect to the index operations in $\IndexOpSet$:
\begin{mathpar}
  \inferrule*
  {(\Delta \vdash e \stackrel{ax}\equiv e' : s) \in \IndexAxiomSet \\
    \sigma : \Delta' \Rightarrow \Delta}
  {\Delta' \vdash \sigma^*e \equiv \sigma^*e' : s}

  \inferrule*
  {\{\Delta \vdash e_j \equiv e'_j : s_j\}_{1\leq j\leq n}}
  {\Delta \vdash \indexOp{f}(e_1, ..., e_n) \equiv \indexOp{f}(e'_1, ..., e'_n) : s}
\end{mathpar}
We also assume the standard reflexivity, symmetry and transitivity
rules for equality.

@@

The judgmental equalities on index terms and on types are preserved by
substitution:
\begin{lemma}
  Let $\sigma : \Delta \Rightarrow \Delta'$ be a simultaneous
  substitution.
  \begin{enumerate}
  \item If $\Delta' \vdash e \equiv e' : s$ then $\Delta \vdash
    \sigma^*e \equiv \sigma^*e' : s$; and
  \item If $\Delta' \vdash A \equiv B \isType$ then $\Delta \vdash
    \sigma^*A \equiv \sigma^* B \isType$.
  \end{enumerate}
\end{lemma}

>>>>

  \sigma^*(\forall i\mathord:s.A) \isDefinedAs \forall i\mathord:s.\sigma_{i\mathord:s}^*A
\end{mathpar}
\begin{lemma}
  Let $\Delta \vdash \sigma \Rightarrow \Delta'$ be a simultaneous
  substitution.
  \begin{enumerate}
  \item If $\Delta' \vdash e : s$, then $\Delta \vdash \sigma^*e : s$; and
  \item If $\Delta' \vdash A \isType$, then $\Delta \vdash \sigma^*A
    \isType$.
  \end{enumerate}
\end{lemma}

@@

\fixme{Consider updating this example} Much of the power of indexing
types by the terms of an algebraic theory comes from the equations of
the theory, and the fact that these equations are inherited by the
types. In the units of measure example, the well-indexed types
$\tyPrim{X}{u_1\cdot u_2}$ and $\tyPrim{X}{u_2 \cdot u_1}$ are
considered equal by the type system, due to the commutativity of the
$-\cdot-$ operation in the theory of abelian groups.

In the general framework, the equations between types are derived from
a set $\IndexAxiomSet$ of axioms $\Delta \vdash e \stackrel{ax}\equiv
e' : s$. We assume that all axioms $(\Delta \vdash e
\stackrel{ax}\equiv e' : s) \in \IndexAxiomSet$ are well-sorted:
meaning that both $\Delta \vdash e : s$ and $\Delta \vdash e' : s$.

Given a set $\IndexAxiomSet$ of axioms, we generate the judgmental
equality between index terms $\Delta \vdash e \equiv e' : s$ by the
following rules, which allow us to use substitution instances of
axioms, and ensure that judgmental equality is a congruence relation
with respect to the index operations in $\IndexOpSet$:
\begin{mathpar}
  \inferrule*
  {(\Delta' \vdash e \stackrel{ax}\equiv e' : s) \in \IndexAxiomSet \\
    \Delta \vdash \sigma \Rightarrow \Delta'}
  {\Delta \vdash \sigma^*e \equiv \sigma^*e' : s}

  \inferrule*
  {\{\Delta \vdash e_j \equiv e'_j : s_j\}_{1\leq j\leq n}}
  {\Delta \vdash \indexOp{f}(e_1, ..., e_n) \equiv \indexOp{f}(e'_1, ..., e'_n) : s}
\end{mathpar}
We also assume the standard reflexivity, symmetry and transitivity
rules for the judgmental equality.

\begin{example}[Origin Invariance]
  FIXME
\end{example}

@@

The judgmental equalities on index terms and on types are preserved by
substitution:
\begin{lemma}
  Let $\Delta \vdash \sigma \Rightarrow \Delta'$ be a simultaneous
  substitution.
  \begin{enumerate}
  \item If $\Delta' \vdash e \equiv e' : s$ then $\Delta \vdash
    \sigma^*e \equiv \sigma^*e' : s$; and
  \item If $\Delta' \vdash A \equiv B \isType$ then $\Delta \vdash
    \sigma^*A \equiv \sigma^* B \isType$.
  \end{enumerate}
\end{lemma}
